---
title: "VIR-CLASP Overlaps"
output: html_notebook
---

The goal of this R Notebook is to determine the proteins that are shared and distinct between the different conditions of VIR-CLASP

#The secondary goal is to determine the proteins that are shared and distinct between these external datasets:
  - The set of Interferon Stimulated Genes (ISGs) which can be found in (filename)
  - The set of hits from the CHIKV siRNA screen which can be found in (filename)
  - The set of RNA binding proteins which can be found in (Gerstberger_2014_tableS3.csv)

#The files used will be the output of the VIR_CLASP_Protein_Filtering Notebook
  - all_proteins.csv
  - high_confidence_vir_clasp.csv
  - medium_confidence_vir_clasp.csv
  
```{r}
library(venneuler)
library(tidyverse)
library(purrr)
library(knitr)

input_directory <- "/Users/saraharcos/Desktop/VIR_CLASP_Analysis/Output/"
output_directory <- "/Users/saraharcos/Desktop/VIR_CLASP_Analysis/Overlaps/"
external_directory <- "/Users/saraharcos/Desktop/VIR_CLASP_Analysis/External_Data/"

```

```{r}
#Function Definitions

clean_names <- function(input_names) {
  id <- gsub("\\w*_HUMAN","",input_names)
  id <- gsub("gi\\|\\w*,", "", id)
  id <- gsub("r-tr\\|\\w*\\|,", "", id)
  id <- gsub("tr\\|\\w*\\|,", "", id)
  id <- gsub(",tr\\|\\w*\\|", "", id)
  id <- gsub(",r-tr\\|\\w*\\|", "", id)
  id <- gsub(",sp\\|\\w*\\|", "", id)
  id <- gsub("r-sp\\|", "", id)
  id <- gsub("sp\\|", "", id)
  id <- gsub("r-tr\\|", "", id)
  id <- gsub("tr\\|", "", id)
  id <- gsub("\\|", "", id)
  id <- data_frame(Proteins = id)
  #return
  colnames(id) <- " "
  id
}
```


##Read in files
```{r message = FALSE}
vir_clasp <- read_csv(paste(input_directory, "all_virclasp.csv", sep = ""))
high_confidence_vir_clasp <- read_csv(paste(input_directory, "high_confidence_vir_clasp.csv", sep = ""))
medium_confidence_vir_clasp <- read_csv(paste(input_directory, "medium_confidence_vir_clasp.csv", sep = ""))
no4SU <- read_csv(paste(input_directory, "no4SU.csv", sep = ""))


```


##Overlap between -4SU and +4SU samples
```{r}
library(venneuler)
vir_clasp_minus <- vir_clasp %>%
  filter(grepl("minus", condition))

no4SU_minus <- no4SU %>%
  filter(grepl("minus", condition))

high_conf_minus <- high_confidence_vir_clasp %>%
  filter(grepl("minus", condition))

n_virclasp_minus <- n_distinct(vir_clasp_minus$Accession)

n_high_conf_minus <- n_distinct(high_conf_minus$Accession)
n_no4SU_minus <- n_distinct(no4SU_minus$Accession)

overlap_minus <- venneuler(c("VIRCLASP" = n_high_conf_minus, 
                            "-4SU" = (n_no4SU_minus - (n_virclasp_minus - n_high_conf_minus)), 
                            "VIRCLASP&-4SU" = (n_virclasp_minus - n_high_conf_minus)))
pdf(paste(output_directory, 'overlap_no4SU_minus.pdf', sep = ""))
plot(overlap_minus)
dev.off()

vir_clasp_plus <- vir_clasp %>%
  filter(grepl("plus", condition))

no4SU_plus <- no4SU %>%
  filter(grepl("plus", condition))

high_conf_plus <- high_confidence_vir_clasp %>%
  filter(grepl("plus", condition))

n_virclasp_plus <- n_distinct(vir_clasp_plus$Accession)
n_high_conf_plus <- n_distinct(high_conf_plus$Accession)
n_no4SU_plus <- n_distinct(no4SU_plus$Accession)

overlap_plus <- venneuler(c("VIRCLASP" = n_high_conf_plus, 
                            "-4SU" = (n_no4SU_plus - (n_virclasp_plus - n_high_conf_plus)), 
                            "VIRCLASP&-4SU" = (n_virclasp_plus - n_high_conf_plus)))
pdf(paste(output_directory, 'overlap_no4SU_plus.pdf', sep = ""))
plot(overlap_plus)
dev.off()
```

##Overlap between VIR-CLASP conditions

```{r}
library(UpSetR)

upset_frame <- high_confidence_vir_clasp %>%
  select(condition, Accession, Name, Description) %>%
  mutate(exists = 1) %>%
  group_by(Accession) %>%
  spread(condition, exists, fill = 0)

minus_IFN <- upset_frame %>%
  filter(hour0_minus == 0 &
         hour1_minus == 0 &
         hour3_minus == 0 &
         hour0_plus == 1 &
         hour1_plus == 0 &
         hour3_plus == 0)


inputData <- as.data.frame(upset_frame)
inputData$Accession <- as.factor(inputData$Accession)
str(inputData)

#write to csv for external use in UpSet
write_delim(inputData, paste(output_directory, "CHIKV_UpSet.csv", sep = ""), delim = ";")

#pass as vector
upset_fxn <- function(datasets){
 upset(inputData, nsets = 6, 
      order.by = c("freq"), 
      sets = datasets,
      queries = list(list(query = intersects, params = list("hour0_plus", "hour1_plus", "hour3_plus"),
                          color = "#FEB938", active = T),
                     list(query = intersects, params = list("hour0_minus", "hour1_minus", "hour3_minus"),
                          color = "blue", active = T)),
      keep.order = TRUE,
      #group.by = "sets",
      cutoff = 3,
      empty.intersections = "on",
      text.scale = .8,
      sets.bar.color = "#56B4E9",
      mainbar.y.label = "Protein Intersections",
      sets.x.label = "Proteins Per Condition",
      line.size = 0.5,
      point.size = 1,
      mb.ratio = c(0.65, 0.35))
}
pdf(paste(output_directory, 'VIR_CLASP_UpSet.pdf', sep = ""), width = 6, height = 2)
upset_fxn(c("hour0_minus", "hour1_minus", "hour3_minus", "hour0_plus", "hour1_plus", "hour3_plus"))
dev.off()

#ggsave(path = output_directory, filename = "VIR_CLASP_UpSet.pdf", plot = r, device = "pdf", height = 5, width = 5)


```

##Gerstberger RBP list
  - Need to tranlate the Uniprot ID's from the VIR-CLASP data to Ensembl Protein ID's
  - See rtf in "External Data" folder for explanation of Gerstberger dataset

```{r}
gerstberger <- read_csv(paste(external_directory, "Gerstberger_Uniprot.csv", sep = "")) %>%
  filter(Status == "reviewed") %>%
  select("Accession" = Entry)
n_distinct(gerstberger$Accession)
```

##GO RNA-binding annotations
  -Retrieved from geneontology.org on 05/03/18

```{r}
go_rbp <- read_tsv(paste(external_directory,"RNA_binding_GO_050318.txt", sep = ""), col_names = c("Accession", "Name"))
go_rbp$Accession <- gsub(pattern = "UniProtKB\\:", go_rbp$Accession, replacement = "")
go_rbp$Name = NULL
n_distinct(go_rbp$Accession)

```

## Castello Interactome
  - Same protocol as Gerstberger, Ensembl ID's were converted to current version, then converted to Uniprot ID`s for comparison
```{r}
castello <- read_csv(paste(external_directory, "HeLaInteractome_uniprot.csv", sep = ""), col_names = "Accession")
n_distinct(castello$Accession)

```

## Baltz Interactome

```{r}
baltz <- read_csv(paste(external_directory, "Baltz_uniprot.csv", sep = "")) %>%
  select(Accession = Entry)
```


##Comparisons between Vir-CLASP High Confidence, GO Annotated RBP's, Gerstberger RBP's, Castello RBP's, and Baltz RBP's

```{r}
#Get Cleaned Uniprot Accessions of VIR-CLASP data
n_distinct(high_confidence_vir_clasp$Accession)

high_confidence_vir_clasp_annotated <- high_confidence_vir_clasp %>%
  filter(Accession %in% go_rbp$Accession |
         Accession %in% gerstberger$Accession |
         Accession %in% castello$Accession |
         Accession %in% baltz$Accession)
n_distinct(high_confidence_vir_clasp_annotated$Accession)

vir_only <- filter(high_confidence_vir_clasp, !(Accession %in% go_rbp$Accession) &
                                                    !(Accession %in% gerstberger$Accession) &
                                                    !(Accession %in% castello$Accession &
                                                    !(Accession %in% baltz$Accession)))
n_distinct(vir_only$Accession)
#Make Table of overlaps

gerstberger2 <- mutate(gerstberger, Gerstberger2014 = "X")
go_rbp2 <- mutate(go_rbp, `GO: RNA-binding` = "X")
castello2 <- mutate(castello, Castello2012 = "X")
baltz2 <- mutate(baltz, Baltz2012 = "X")
annotated_rbps <- high_confidence_vir_clasp_annotated %>%
  select(-condition, -ave_spectra) %>%
  unique() %>%
  full_join_NA(gerstberger2, by = "Accession") %>%
  full_join_NA(castello2, by = "Accession") %>%
  full_join_NA(go_rbp2, by = "Accession") %>%
  full_join_NA(baltz2, by = "Accession") %>%
  filter(Name != 0)

write_csv(annotated_rbps, paste(output_directory, "Sup_Table_Annotated_RBPs.csv", sep = ""))

```

##siRNA Screen for CHIKV (Karlas et al 2016, Nature Communications)
 - Use combined high and medium confidence vir-clasp sets

```{r}
n_distinct(medium_confidence_vir_clasp$Accession)


all_confidence_vir_clasp <- rbind(high_confidence_vir_clasp, medium_confidence_vir_clasp)
n_distinct(all_confidence_vir_clasp$Accession)

karlas_proviral <- read_csv(paste(external_directory, "Karlas_proviral_uniprot.csv", sep = ""), col_names = "Accession")
n_distinct(karlas_proviral$Accession)

proviral_vir_clasp <- all_confidence_vir_clasp %>%
  filter(Accession %in% karlas_proviral$Accession) %>%
  mutate(CHIKV_Effect = "Pro-Viral") %>%
  select(-ave_spectra)
n_distinct(proviral_vir_clasp$Accession)

karlas_antiviral <- read_csv(paste(external_directory, "Karlas_antiviral_uniprot.csv", sep = ""), col_names = "Accession")
n_distinct(karlas_antiviral$Accession)

antiviral_vir_clasp <- all_confidence_vir_clasp %>%
  filter(Accession %in% karlas_antiviral$Accession) %>%
  mutate(CHIKV_Effect = "Anti-Viral") %>%
  select(-ave_spectra)
n_distinct(antiviral_vir_clasp$Accession)

full_karlas <- rbind(proviral_vir_clasp, antiviral_vir_clasp)
full_karlas <- full_karlas %>%
  mutate(exists = "X") %>%
  spread(condition, exists, fill = "") %>%
  arrange(CHIKV_Effect)
write_csv(full_karlas, path = paste(output_directory, "Sup_Table_Karlas_screen.csv", sep = ""))


```

##Overlap between plus and minus IFN

```{r}
minus_vir_clasp <- high_confidence_vir_clasp %>%
  filter(condition %in% c("hour0_minus", "hour1_minus", "hour3_minus"))

plus_vir_clasp <- high_confidence_vir_clasp %>%
  filter(condition %in% c("hour0_plus", "hour1_plus", "hour3_plus"))

minus_and_plus <- minus_vir_clasp %>%
  filter(Accession %in% plus_vir_clasp$Accession)

minus_only <- minus_vir_clasp %>%
  filter(!(Accession %in% plus_vir_clasp$Accession))

plus_only <- plus_vir_clasp %>%
  filter(!(Accession %in% minus_vir_clasp$Accession))

minus <- n_distinct(minus_vir_clasp$Accession)
plus <- n_distinct(plus_vir_clasp$Accession)
both <- n_distinct(minus_and_plus$Accession)
minus_o <- n_distinct(minus_only$Accession)
plus_o <- n_distinct(plus_only$Accession)


overlap_minus_plus <- venneuler(c("Plus" = plus_o, 
                            "Minus" = minus_o, 
                            "Plus&Minus" = both))
pdf(paste(output_directory, 'overlap_minus_plus.pdf', sep = ""))
plot(overlap_minus_plus)
dev.off()
```

##Output the minus and plus only lists for GO analysis

```{r}
write_csv(minus_only, path = paste(output_directory, "high_confidence_minus_only.csv", sep = ""))
write_csv(plus_only, path = paste(output_directory, "high_confidence_plus_only.csv", sep = ""))
```


##Output the minus and plus only lists for supplementary tables
```{r}
n_distinct(minus_only$Accession)
n_distinct(plus_only$Accession)

minus_only_sup <- minus_only %>%
  mutate(exists = "X") %>%
  select(-ave_spectra) %>%
  spread(condition, exists, fill = "")

plus_only_sup <- plus_only %>%
  mutate(exists = "X") %>%
  select(-ave_spectra) %>%
  spread(condition, exists, fill = "")

write_csv(minus_only_sup, path = paste(output_directory, "Sup_Table_minus_only.csv", sep = ""))
write_csv(plus_only_sup, path = paste(output_directory, "Sup_Table_plus_only.csv", sep = ""))
```

